---
title: "ppEffect Report"
author: "YAOJ"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: true 
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(cowplot)
library(dplyr)
library(plotly)
library(ggvenn)
library(RColorBrewer)
library(reshape2)
```

# Overview of the dataset
```{r echo=FALSE}
Up_ppGroup_Q2 <- object@meta.data %>% select(group.by,Up_ppDEGs) %>% group_by(get(group.by)) %>% 
  summarize(group.by, median = median(Up_ppDEGs)) %>% filter(median>=median(object$Up_ppDEGs)) %>% as.data.frame() 
Up_ppGroup_Q2 <- droplevels ( Up_ppGroup_Q2$`get(group.by)`)

Up_ppGroup_Q3 <- object@meta.data %>% select(group.by,Up_ppDEGs) %>% group_by(get(group.by)) %>% 
  summarize(group.by, median = median(Up_ppDEGs)) %>% filter(median>=quantile(object$Up_ppDEGs,0.75)) %>% as.data.frame() 
Up_ppGroup_Q3 <- droplevels ( Up_ppGroup_Q3$`get(group.by)`)




Down_ppGroup_Q2 <- object@meta.data %>% select(group.by,Down_ppDEGs) %>% group_by(get(group.by)) %>% 
  summarize(group.by, median = median(Down_ppDEGs)) %>% filter(median<=median(object$Down_ppDEGs)) %>% as.data.frame() 
Down_ppGroup_Q2 <- droplevels ( Down_ppGroup_Q2$`get(group.by)`)

Down_ppGroup_Q1 <- object@meta.data %>% select(group.by,Down_ppDEGs) %>% group_by(get(group.by)) %>% 
  summarize(group.by, median = median(Down_ppDEGs)) %>% filter(median<=quantile(object$Down_ppDEGs,0.25)) %>% as.data.frame() 
Down_ppGroup_Q1 <- droplevels ( Down_ppGroup_Q1$`get(group.by)`)


```

+---------------------------------------+---------------------------------+
|  Attribute                            |     Content                     | 
+=======================================+=================================+
|  Object name                          | `r object@project.name`         |
+---------------------------------------+---------------------------------+
|  Cell number                          | `r length(colnames(object))`    |
+---------------------------------------+---------------------------------+
|  Gene number                          | `r length(rownames(object))`    |
+---------------------------------------+---------------------------------+
|  Active assay                         | `r object@active.assay`         |
+---------------------------------------+---------------------------------+
|  Reductions                           | `r names(object@reductions)`    |
+---------------------------------------+---------------------------------+
| Up-regulated pp.Score (Q2)            | `r median(object$Up_ppDEGs)`    |
+---------------------------------------+---------------------------------+
| Up-regulated Groups in terms of Q2    | `r paste(Up_ppGroup_Q2,' ')`    |
+---------------------------------------+---------------------------------+
| Up-regulated pp.Score (Q3)            | `r quantile(object$Up_ppDEGs,0.75)`|
+---------------------------------------+---------------------------------+
| Up-regulated Groups in terms of Q3    | `r paste(Up_ppGroup_Q3,' ')`    |
+---------------------------------------+---------------------------------+
| Down-regulated pp.Score (Q2)            | `r median(object$Down_ppDEGs)`    |
+---------------------------------------+---------------------------------+
| Down-regulated Groups in terms of Q2    | `r paste(Down_ppGroup_Q2,' ')`    |
+---------------------------------------+---------------------------------+
| Down-regulated pp.Score (Q1)            | `r quantile(object$Down_ppDEGs,0.75)`|
+---------------------------------------+---------------------------------+
| Down-regulated Groups in terms of Q1    | `r paste(Down_ppGroup_Q1,' ')`    |
+---------------------------------------+---------------------------------+

> **Gene number**: representing the gene numbers in the activate assay.
> 
> **Reductions**: representing the dimension reduction methods conducted in this dataset.
> 
> **Up-regulated pp.Score (Q2)** : representing the median expression value of enzymolysis induced genes (Up_ppDEGs).
> 
> **Up-regulated Groups in terms of Q2** : calculating the median of pp.Score in each cell group, then to figure out which clusters have higher pp.Scores than the pp.Score (Q2).
> 
> **Up-regulated pp.Score (Q3)** : representing The third quartile(Q3), also known as the higher quartile, is equal to the 75% of all the values in the sample arranged from the smallest to the largest of enzymolysis induced genes (Up_ppDEGs).
> 
> **Up-regulated Groups in terms of Q3** : calculating the median of pp.Score in each cell group, then to figure out which clusters have higher pp.Scores than the pp.Score (Q3).

> **Down-regulated pp.Score (Q2)** : representing the median expression value of enzymolysis induced genes (Down_ppDEGs).
> 
> **Down-regulated Groups in terms of Q2** : calculating the median of pp.Score in each cell group, then to figure out which clusters have lower pp.Scores than the pp.Score (Q2).
> 
> **Down-regulated pp.Score (Q1)** : representing The first quartile(Q1), also known as the lower quartile, is equal to the smallest 25% of all the pp.Score values of enzymolysis induced genes (Down_ppDEGs).
> 
> **Down-regulated Groups in terms of Q1** : calculating the median of pp.Score in each cell group, then to figure out which clusters have lower pp.Scores than the pp.Score (Q1).


# Overlap between ppDEGs and HVGs

In this dataset, we totally have <span style="text-decoration: underline;background-color:yellow;">`r length(VariableFeatures(object = object))`</span> highly variable genes (HVGs), and all the HVGs will be used for PCA analysis following up. 

The number of **Up_ppDEGs** is <span style="text-decoration: underline;background-color:yellow;">`r length(Up_ppDEGs)`</span>.
The intersection between HVGs and Up_ppDEGs were <span style="text-decoration: underline;background-color:yellow;">`r length(intersect(VariableFeatures(object = object),Up_ppDEGs ))`</span>  (genes),
which means that  <span style="text-decoration: underline;background-color:yellow;">  `r paste0(round(length(intersect(VariableFeatures(object = object),Up_ppDEGs ))*100/(length(Up_ppDEGs)),2),"%")`</span> Up_ppDEGs belong to HVGs. 

The number of **Down_ppDEGs** is <span style="text-decoration: underline;background-color:yellow;">`r length(Down_ppDEGs)`</span>.The intersection between HVGs and Down_ppDEGs were <span style="text-decoration: underline;background-color:yellow;">`r length(intersect(VariableFeatures(object = object),Down_ppDEGs ))`</span>  (genes),
which means that  <span style="text-decoration: underline;background-color:yellow;">  `r paste0(round(length(intersect(VariableFeatures(object = object),Down_ppDEGs ))*100/(length(Down_ppDEGs)),2),"%")`</span> Down_ppDEGs belong to HVGs.


```{r echo=FALSE, warning=TRUE}
if (length(VariableFeatures(object = object))<=100){
  print("please re-run the funtion FindVariableFeatures on your object first")
} else {
  HVGs <-VariableFeatures(object = object)
  plot <- ggvenn(list( "Up_ppDEGs"= Up_ppDEGs, "Down_ppDEGs"= Down_ppDEGs,"HVGs" = HVGs), c("Up_ppDEGs","Down_ppDEGs","HVGs"),fill_alpha =0.3, fill_color = selCols[c(2,3,1)])
  ggplotly(plot)
}
```



# The extent of ppEffects in each cell
In this panel, your can see the distribution of ppEffects in each cells by UMAP plot. The darker the cell color, the more severely it is affected by enzymatic hydrolysis



```{r tsne, echo=F, fig.height=15, fig.width=12, warning=F, dev='svg'}
plot1 <- Seurat::DimPlot(object= object, group.by = group.by, cols = selCols,reduction = "umap",pt.size = 0.5,label= T, label.box = T )
plot2 <- Seurat::DimPlot(object= object, group.by = group.by, cols = selCols,reduction = "tsne",pt.size = 0.5, label = T, label.box = T)

plot3 <- Seurat::FeaturePlot(object = object, reduction = "umap",features = "Up_ppDEGs",cols =featurePlotCols,pt.size = 0.5)
plot4 <- Seurat::FeaturePlot(object = object, reduction = "tsne",features = "Up_ppDEGs",cols =featurePlotCols,pt.size = 0.5)

plot5 <- Seurat::FeaturePlot(object = object, reduction = "umap",features = "Down_ppDEGs",
                             cols =rev( brewer.pal(9, "Blues")),pt.size = 0.5)
plot6 <- Seurat::FeaturePlot(object = object, reduction = "tsne",features = "Down_ppDEGs",
                             cols =rev( brewer.pal(9, "Blues")),pt.size = 0.5)

cowplot::plot_grid(plot1, plot2 ,plot3,plot4 ,plot5,plot6,ncol = 2)
```



# Groups affected by ppEffect 

In this panel, we use the violin plot to  display which groups were highly affected by ppEffect.

## Up-regulated groups

>The ***red dotted line*** presented the median expression value of enzymolysis induced genes (Up_ppDEGs). The median expression value of Up_ppDEGs (Up-regulated pp.Score Q2) ***in this dataset*** is <span style="text-decoration: underline;background-color:yellow;"> `r median(object$Up_ppDEGs)`</span> .
>
>The ***yellow dotted line*** presented the higher quartile expression value of enzymolysis induced genes (Up_ppDEGs). The higher quartile expression value of Up_ppDEGs (Up-regulated pp.Score Q3) ***in this dataset*** is <span style="text-decoration: underline;background-color:yellow;">`r quantile(object$Up_ppDEGs,0.75)`</span> .

```{r VlnPlot, echo=FALSE,warning=FALSE,fig.align="center", fig.height=4, fig.width=8,dev='svg'}
p <- VlnPlot(object,features = "Up_ppDEGs",cols = selCols)+
  geom_hline(aes(yintercept= median(object$Up_ppDEGs)),colour ="#990000", linetype ="dashed",size = 1)+
  geom_hline(aes(yintercept= quantile(object$Up_ppDEGs,0.75)),colour ="yellow", linetype ="dashed",size = 1)
 ggplotly(p) 
```
In this panel, we use the Dotplot to display which groups were highly affected by Up_ppDEGs. 

```{r Dotplot, echo=FALSE,fig.align="center", fig.height=3, fig.width=8,dev='svg'}
p <- DotPlot(object,features = "Up_ppDEGs",cols = "RdGy")+coord_flip()+
  ggtitle("Scores of Up_ppDEGs in each groups")
ggplotly(p)
```


## Down-regulated groups

>The ***red dotted line*** presented the median expression value of enzymolysis induced genes (Down_ppDEGs). The median expression value of Down_ppDEGs (Down-regulated pp.Score Q2) ***in this dataset*** is <span style="text-decoration: underline;background-color:yellow;"> `r median(object$Down_ppDEGs)`</span> .
>
>The ***yellow dotted line*** presented the higher quartile expression value of enzymolysis induced genes (Down_ppDEGs). The lower quartile expression value of Down_ppDEGs (Down-regulated pp.Score Q1) ***in this dataset*** is <span style="text-decoration: underline;background-color:yellow;">`r quantile(object$Down_ppDEGs,0.25)`</span> .

```{r VlnPlot2, echo=FALSE,fig.align="center", fig.height=4, fig.width=8,dev='svg'}
p <- VlnPlot(object,features = "Down_ppDEGs",cols = selCols)+
  geom_hline(aes(yintercept= median(object$Down_ppDEGs)),colour ="#990000", linetype ="dashed",size = 1)+
  geom_hline(aes(yintercept= quantile(object$Down_ppDEGs,0.25)),colour ="yellow", linetype ="dashed",size = 1)
 ggplotly(p) 
```
In this panel, we use the Dotplot to display which groups were highly affected by Down_ppDEGs. 
```{r Dotplot2, warning=FALSE,echo=FALSE,fig.align="center", fig.height=3, fig.width=8,dev='svg'}
p <- DotPlot(object,features = "Down_ppDEGs")+coord_flip()+
  ggtitle("Scores of Down_ppDEGs in each groups")
ggplotly(p)
```


# Overlap between ppDEGs and cluster markers

This panel display the intersection between Up_ppDEGs and marker genes cluster by cluster. 
```{r echo=FALSE, fig.height=4, fig.width=8, warning=FALSE,dev='svg'}

if (!exists("marker_genes")){
  print("Please provide a data.frame from the parameter'marker_genes', you can produce it by the function FindAllMarkers!")
}else {
  # 计算每个cluster中gene的数量
  marker_genes_count <- marker_genes %>% group_by(cluster) %>% summarize(count = n())

  # 计算每个cluster中同时属于Up_ppDEGs的gene数量
  Up_ppDEGs_count <- marker_genes %>% filter(gene %in% Up_ppDEGs) %>% group_by(cluster) %>% summarize(count = n())

  # 合并两个数据集
  df <- merge(marker_genes_count, Up_ppDEGs_count, by = 'cluster', all.x = TRUE)

  # 堆叠柱状图
  df_melt <- melt(df, id.vars = 'cluster')
  df_melt$variable <- gsub("count.y","Up_ppDEGs",df_melt$variable)
  df_melt$variable <- gsub("count.x","Markers",df_melt$variable)
  p1 <- ggplot(df_melt, aes(x = cluster, y = value, fill = variable)) + geom_bar(stat = 'identity', position = 'stack')+
    theme_bw()+ scale_fill_manual(values=c("#DCB717","#E2272E"))+labs(color = "Gene Number")
  ggplotly(p1)  

}

```


This panel display the intersection between Down_ppDEGs and marker genes cluster by cluster. 
```{r echo=FALSE, fig.height=4, fig.width=8, warning=FALSE,dev='svg'}

if (!exists("marker_genes")){
  print("Please provide a data.frame from the parameter'marker_genes', you can produce it by the function FindAllMarkers!")
}else {
  # 计算每个cluster中gene的数量
  marker_genes_count <- marker_genes %>% group_by(cluster) %>% summarize(count = n())

  # 计算每个cluster中同时属于Down_ppDEGs的gene数量
  Down_ppDEGs_count <- marker_genes %>% filter(gene %in% Down_ppDEGs) %>% group_by(cluster) %>% summarize(count = n())

  # 合并两个数据集
  df <- merge(marker_genes_count, Down_ppDEGs_count, by = 'cluster', all.x = TRUE)

  # 堆叠柱状图
  df_melt <- melt(df, id.vars = 'cluster')
  df_melt$variable <- gsub("count.x","Markers",df_melt$variable)
  df_melt$variable <- gsub("count.y","Down_ppDEGs",df_melt$variable)
  df_melt$variable <- factor(df_melt$variable , levels = c("Markers","Down_ppDEGs"))
  p1 <- ggplot(df_melt, aes(x = cluster, y = value, fill = variable)) + geom_bar(stat = 'identity', position = 'stack')+
    theme_bw()+scale_fill_manual(values=c("#DCB717","#5EA5C9"))
  ggplotly(p1)  

}

```







