---
title: "ppEffect Report"
author: "YAOJ"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: true 
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(cowplot)
library(dplyr)
library(plotly)
library(ggvenn)

```

# Overview of the dataset
```{r echo=FALSE}
ppGroup_Q2 <- object@meta.data %>% select(group.by,ppDEGs) %>% group_by(get(group.by)) %>% 
  summarize(group.by, median = median(ppDEGs)) %>% filter(median>=median(object$ppDEGs)) %>% as.data.frame() 
ppGroup_Q2 <- droplevels ( ppGroup_Q2$`get(group.by)`)

ppGroup_Q3 <- object@meta.data %>% select(group.by,ppDEGs) %>% group_by(get(group.by)) %>% 
  summarize(group.by, median = median(ppDEGs)) %>% filter(median>=quantile(object$ppDEGs,0.75)) %>% as.data.frame() 
ppGroup_Q3 <- droplevels ( ppGroup_Q3$`get(group.by)`)
```

+-------------------------+---------------------------------+
|  Attribute              |     Content                     | 
+=========================+=================================+
|  Object name            | `r object@project.name`         |
+-------------------------+---------------------------------+
|  Cell number            | `r length(colnames(object))`    |
+-------------------------+---------------------------------+
|  Gene number            | `r length(rownames(object))`    |
+-------------------------+---------------------------------+
|  Active assay           | `r object@active.assay`         |
+-------------------------+---------------------------------+
|  Reductions             | `r names(object@reductions)`    |
+-------------------------+---------------------------------+
|pp.Score (Q2)            | `r median(object$ppDEGs)`       |
+-------------------------+---------------------------------+
|ppGroups in terms of Q2  | `r paste(ppGroup_Q2,' ')`       |
+-------------------------+---------------------------------+
|pp.Score (Q3)            | `r quantile(object$ppDEGs,0.75)`|
+-------------------------+---------------------------------+
|ppGroups in terms of Q3  | `r paste(ppGroup_Q3,' ')`       |
+-------------------------+---------------------------------+


> **Gene number**: representing the gene numbers in the activate assay.
> 
> **Reductions**: representing the dimension reduction methods conducted in this dataset.
> 
> **pp.Score (Q2)** : representing the median expression value of enzymolysis induced genes (ppDEGs).
> 
> **ppGroups in terms of Q2** : calculating the median of pp.Score in each cell group, then to figure out which clusters have higher pp.Scores than the pp.Score (Q2).
> 
> **pp.Score (Q3)** : representing The third quartile(Q3), also known as the higher quartile, is equal to the 75% of all the values in the sample arranged from the smallest to the largest of enzymolysis induced genes (ppDEGs).
> 
> **ppGroups in terms of Q3** : calculating the median of pp.Score in each cell group, then to figure out which clusters have higher pp.Scores than the pp.Score (Q3).


# Overlap between ppDEGs and HVGs

In this dataset, we totally have <span style="text-decoration: underline;background-color:yellow;">`r length(VariableFeatures(object = object))`</span> highly variable genes (HVGs),
and the number of ppDEGs is <span style="text-decoration: underline;background-color:yellow;">`r length(ppDEGs)`</span>. 
The intersection between these two sets were <span style="text-decoration: underline;background-color:yellow;">`r length(intersect(VariableFeatures(object = object),ppDEGs ))`</span>  (genes),
which means that  <span style="text-decoration: underline;background-color:yellow;">  `r paste0(round(length(intersect(VariableFeatures(object = object),ppDEGs ))*100/(length(ppDEGs)),2),"%")`</span> ppDEGs belong to HVGs. All the HVGs will be used for PCA analysis following up. 


```{r echo=FALSE, warning=TRUE}
if (length(VariableFeatures(object = object))<=100){
  print("please re-run the funtion FindVariableFeatures on your object first")
} else {
  HVGs <-VariableFeatures(object = object)
  plot <- ggvenn(list("HVGs" = HVGs, "ppDEGs"= ppDEGs), c("HVGs","ppDEGs"),fill_alpha =0.3, fill_color = selCols[1:2])
  ggplotly(plot)
}

```




# The extent of ppEffects in each cell
In this panel, your can see the distribution of ppEffects in each cells by UMAP plot. The darker the cell color, the more severely it is affected by enzymatic hydrolysis


```{r tsne, echo=F, fig.height=10, fig.width=12, warning=F, dev='svg'}
plot1 <- Seurat::FeaturePlot(object = object, features = "ppDEGs",cols =featurePlotCols,pt.size = 0.8 )
plot2 <- Seurat::DimPlot(object= object, group.by = group.by, cols = selCols,reduction = "umap",pt.size = 0.8 )
plot3 <- Seurat::FeaturePlot(object = object, reduction = "tsne",features = "ppDEGs",cols =featurePlotCols,pt.size = 0.8)
plot4 <- Seurat::DimPlot(object= object, group.by = group.by, cols = selCols,reduction = "tsne",pt.size = 0.8)

cowplot::plot_grid(plot1, plot2 ,plot3,plot4 ,ncol = 2)
```

# Groups affected by ppEffect 

In this panel, we use the violin plot to  display which groups were highly affected by ppEffect.

>The ***red dotted line*** presented the median expression value of enzymolysis induced genes (ppDEGs). The median expression value of ppDEGs (pp.Score Q2) ***in this dataset*** is <span style="text-decoration: underline;background-color:yellow;"> `r median(object$ppDEGs)`</span> .
>
>The ***yellow dotted line*** presented the higher quartile expression value of enzymolysis induced genes (ppDEGs). The higher quartile expression value of ppDEGs (pp.Score Q3) ***in this dataset*** is <span style="text-decoration: underline;background-color:yellow;">`r quantile(object$ppDEGs,0.75)`</span> .

```{r VlnPlot, echo=FALSE,fig.align="center", fig.height=4, fig.width=8,dev='svg'}
p <- VlnPlot(object,features = "ppDEGs",cols = selCols)+
  geom_hline(aes(yintercept= median(object$ppDEGs)),colour ="#990000", linetype ="dashed",size = 1)+
  geom_hline(aes(yintercept= quantile(object$ppDEGs,0.75)),colour ="yellow", linetype ="dashed",size = 1)
 ggplotly(p) 
```


In this panel, we use the Dotplot to display which groups were highly affected by ppEffect. 

```{r Dotplot, echo=FALSE,fig.align="center", fig.height=3, fig.width=8,dev='svg'}
p <- DotPlot(object,features = "ppDEGs",cols = "RdGy")+coord_flip()
ggplotly(p)
```


# Overlap between ppDEGs and cluster markers

This panel display the intersection between ppDEGs and marker genes cluster by cluster. Marker gene list were pro 
```{r echo=FALSE, fig.height=9, fig.width=8, warning=FALSE,dev='svg'}

if (!exists("marker_genes")){
  print("Please provide a data.frame from the parameter'marker_genes', you can produce it by the function FindAllMarkers!")
}else {
  plot_list =list()
for (i in 0: (length(unique(marker_genes$cluster))-1)){
  Markers =filter(marker_genes, cluster==i)[["gene"]]
  cluster_name = paste0("C",i)
  plot_list[[cluster_name]] <- ggvenn(list("markers" = Markers, "ppDEGs"= ppDEGs), c("markers","ppDEGs"),fill_alpha =0.3, fill_color = c("#DCB717","#E2272E"),set_name_size=3,show_percentage=F)+ggtitle(cluster_name) 
}

plot_grid( plotlist = plot_list,ncol = 4)
}

```





